Understanding the technological transition based on cultural transmission biases during the Late Pleistocene in Korea

### 2020-05-26 analysis/paper/paper.Rmd
added several para for cultural evolution

3. What is our conceptual framework?
- Evolutionary concepts that links stone artifacts to human behavior
- Cultural Evolution : the idea that cultural change constitutes an evolutionary process
- Same general concept of biological evolution: genetic change in population that is inherited over generations. Apply this concept to artifacts or cultural phenomena
- In this paper, I will use Cultural Transmission 
- Cultural Transmssion: way that people learn and pass on information
- Transmission biases
- similar study (CT in stone artifacts) : Mesoudi&O'Brien, Bettinger&Eerkens, Garvey

>>>>

3. What is our conceptual framework?
- Evolutionary concepts that links stone artifacts to human behavior
- Cultural Evolution : the idea that cultural change constitutes an evolutionary process
- Same general concept of biological evolution: genetic change in population that is inherited over generations. Apply this concept to artifacts or cultural phenomena

- In this paper, I will use Cultural Transmission 
- Cultural Transmssion: way that people learn and pass on information
- Transmission biases
- similar study (CT in stone artifacts) : Mesoudi&O'Brien, Bettinger&Eerkens, Garvey

-------------
Application the theoretical and methodological tenets of Darwinian evolutionary concpet to archaeological studies has helped understand technological transition related to human behavior (Dunnell 1980; Lipo et al. 1997; Bettinger and Eerkens 1997, 1999, 2008, Mesoudi abd O'Brien 2008).

biological concept to archaeological usage: (Richerson and Boyd 1992)

Cultrual evolution that archaeologists adapted has three sub-categories: human behavioral ecology, phylogenetics and cultural transmission theory (Garvey 2018). Among them, we apply models and methods of cultural tranmssion to understand the technological change and socio cultural contexts.

Cultural transmission is the idea that a transition can occur during the process of social learning and the transmission biases are important loci of changes in material cultural (Lycett 2015; Creanza et al. 2017). In other words, this concept holds that instructions for making artifacts are transmitted between members of a group during social learning by observation of, or interaction with another member (Heyes 1994; Eerkens and Lipo 2007; Brantingham et al. 2010; Lycett 2015).

The transmission is subject to biases that result from different social contexts. Boyd and Richerson (1985) and Henrich and McElreath (2003) have defined many of these biases, including ‘guided variation’ (where individuals learn new behaviors and then modify them through trial and error), ‘content-based bias’ (where some aspect of the transmitted instructions, such as cultural preferences, makes them more likely to be adopted), ‘frequency-based bias’ (where an individual is biased to choose particular instructions based on their perceived frequency in the population, such as extremely popular or rare instructions), and ‘model-based bias’ (where a variant is transmitted because of its association with other attributes, such as the prestige or skill of other individuals).

Bettinger and Eerkens (1997; 1999; 2008) have shown how these biases can be identified in stone artifact assemblages. They studied metric variables of stone points during the introduction of bow and arrow technology in the Great Basin around AD 300-600. They equated guided transmission with high metric variation and low correlation between metric attributes, and equated model-based bias with less variation and more correlated variables. They found that in eastern California artifacts have poor correlations of basal width and mass, suggesting new technology was introduced and spread by guided transmission.On the other hand, in central Nevada metric variables are highly correlated, indicating that the bow and arrow was introduced and spread by model-based bias. Since Bettiniger and Eerkens studied archaeological data so they are unable to directly observe the transmission process. To test the underlying assumptions that Bettinger and Eerkens used, Mesoudi and O’Brien (2008) have conducted experiments that simulated prestige bias and guided variation, and observed results that supported Bettinger and Eerkens’ hypothesis.

Garvey

-------------
